---
title: "Stat 184 Project"
author: "ProjectTeam13-3"
date: "May 2, 2021"
output: html_notebook 
---

#### Load Packages
```{r message = FALSE}
library(tidyverse) 
```

#### Read in Data and Convert Stat files to data frames
```{r}
Memfan <- read.csv("Memorial_Fantasy.csv", header = T)
Pgafan <- read.csv("PGA_Fantasy.csv", header = T)
Schfan <- read.csv("Schwab_Fantasy.csv", header = T)
Memstat<- read.delim("Memorial_Stats.txt", header = T )
Memstats <-  as.data.frame(matrix(Memstat$Data, ncol=9, byrow=TRUE))
Pgastat<- read.delim("PGA_Stats.txt", header = T)
Pgastats <-  as.data.frame(matrix(Pgastat$Data, ncol=9, byrow=TRUE))
Schstat<- read.delim("Schwab_Stats.txt", header = T)
Schstats <-  as.data.frame(matrix(Schstat$Data, ncol=9, byrow=TRUE))
```

#### Join Fantasy and Statistics to Create One Data Frame for Each of Three Tournaments

```{r}
Memorial <- Memstats %>%
  inner_join(Memfan, by=c("V2"= "Name"))
Pga <- Pgastats %>%
  inner_join(Pgafan, by=c("V2"= "Name")) 
Schwab <- Schstats %>%
  inner_join(Schfan, by=c("V2"= "Name"))
```

#### Combine the Three Tournament Data Frames to Final Data Frame
###### During this chunk, we create the final data frame, TournamentsFinal, by using rbind to stack the rows from the three different tables. Next, we created a new data frame, Scores, and use TournamentsFinal to spread where each person has a row with scores from the three tournaments. After this, we needed to create the following variables: rank, top10, madecut, salary, tournplayed, the scores from the three tournaments, and their averages across each of the tournaments they have played. The rank function was used in order to assign a rank to each person. That rank variable was then applied to an if-else in order to add the variable top10, assigning a "Yes" to anyone ranked in the top 10. Madecut uses a grepl with the key "CUT" to filter those who made the cut and those who didn't. Salary uses ifelse functions to determine their salary range. Tournplayed is a simple variable where we used a group_by and a summarise n() to find how many tournaments they played. The scores for each tournament were found earlier in the "Scores" data frame and joined in here. Finally the average is found by using rowMeans over the 3 variables for score from the three tournaments and rounded to 1 decimal. 
```{r}
All <- rbind(Memorial, Pga, Schwab)
names(All)[1:9] <- c('Position', 'Player', 'Score', 
                          'Putt', 'ARG', 'App', 'OTT', 'T2G', 'Total')
All$Score <- fct_recode(All$Score, "0" = "E") 
Top10 <- 
  All %>%
  group_by(Tournament) %>%
  mutate(Score = as.numeric(as.character(factor(Score)))) %>%
  mutate(rank = rank(Score, ties.method = "min")) %>%
  mutate(top10 = if_else(rank <= 10, 'Yes', 'No')) %>%
  mutate(MadeCut = if_else(Position =='CUT', 'No', 'Yes')) %>%
  mutate(SalaryRange = ifelse(grepl("6", Salary), "6000s",
                       ifelse(grepl("7", Salary), "7000s",
                       ifelse(grepl("8", Salary), "8000s",
                       ifelse(grepl("9" , Salary), "9000s",
                       ifelse(grepl("10", Salary), "10000s", "11000s"))))))
TournPlayed <- All %>%
  select((Player)) %>%
  group_by(Player) %>%
  mutate(tournplayed = n()) %>%
  distinct()
AvgScore <- All %>%
    select(Player, Score, Tournament) %>%
    mutate(Score = as.numeric(as.character(factor(Score)))) %>%
    spread(Tournament, Score) %>%
    mutate(avg = round(rowMeans(.[2:4], na.rm=TRUE), 1))
Results_Player <- Top10 %>%
  left_join(TournPlayed, by = "Player")
AllResults <- Results_Player %>%
  left_join(AvgScore, by = "Player") %>%
  mutate(Putt = as.numeric(Putt), 
         OTT = as.numeric(OTT))
```

### Research Questions

#### Question 1
#### Does a higher putting score correlate to an overall lower score when comparing all three tournaments?
###### Since the variables Putt, ARG, App, etc. are the average of strokes gained in that aspect of the game, a higher number in Putt would mean that the players are getting to the green earlier and having to putt more. But does this mean that if a player has a 2 Putt score they will be closer to the top of the leaderboard?

###### As shown below, each of the tournaments have a negative correlation between score and the Putt score. Even though there is a negative correlation does not mean that a player having a high putt score will be close to the top of the leaderboard. This would mainly depend on the tournament, since putting seems to be bigger factors in both the Charles Schwab and PGA Championship tournaments. So, if I was a player that was better at getting to the green and didn't have a higher Putt score I would want to join the Memorial tournament. Since with the data we have, it is the tournament that is least affected by Putt score. There is still a significant negative correlation, but it would give the player a better chance.

```{r}
AllResults %>%
  ggplot(aes(x = Putt, y = Score)) +
  geom_point(size = 1L, colour = "#0c4c8a") +
  theme_minimal() +
  facet_wrap(~Tournament)+
  stat_smooth(method = "lm")
```

#### Question 2
#### Does a higher off-the-tee (OTT) score share a positive correlation with a better performance of the player for each of the three tournaments?

The off-the-tee shot is by far the furthest shot taken on every hole, thus inferring that the distance covered here must be crucial to a player scoring well. As a result, a study was conducted to determine if a relationship exists between the OTT score and each player's overall score in the three tournaments.

In the figures below, each observation plotted is representative for a player's overall score versus their score off-the-tee. In this case of determining a better performance than another, we can define this as a combination of both the overall score which is more negative along with the OTT score which is more positive. As a result of their scoring being inverse of each other, a player who performed "better" on these graphs would be located more to the right and downward of his competitors. As one can observe here, each tournament shows weak signs of correlation. From these three plots, I would conclude that the Memorial Tournament shows the best signs of a negative correlation in this data. Conversely, the Charles Schwab tournament exhibits a slight positive correlation. We can draw some inference that sending the ball further on the initial swing off-the-tee does have a positive impact on the player's overall score, but not by much.
```{r}
ggplot(AllResults, aes(x = OTT, y = Score)) + 
  geom_point(size = 1,
             colour = "Dark Green", 
             na.rm = T) + # Remove obs with missing values
  stat_smooth(method = "lm", se = FALSE, 
              color = 'Green', na.rm = T) + # Add a fitted linear regression line
  theme(plot.title = element_text(hjust = 0.5), 
        plot.subtitle = element_text(hjust = 0.5)) +
  facet_wrap(~Tournament, ncol = 3) +
  labs(title = "Relationship Between Player's \nOTT Score and Overall Performance",
       subtitle = "Better performance can be identified closest to the bottom right",
       x = "Off-the-Tee (OTT Score)", 
       y = "Score")

# Save the graph as a PNG file
ggsave("OTTScores.png",
       device = "png",
       width = 8,
       height = 3.5,
       units = "in",
       dpi = 350)
```
